Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes

### 2022-11-17 Manuscript_rev1.tex
Add files via upload

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
Section~\ref{sec:Bayes} covered how Bayes theorem can be applied to estimate the orientation of a dimer based on its scattering signal. To evaluate the estimation we can use the simulation in Section~\ref{sec:brownian}; we denote the dimer's \emph{actual} orientation as $\hat{\bf s}$ and use MSTM to calculate its light scattering $I(\hat{\bf s}, \theta)$. At each angle $\theta_k$ we use \eqref{eq:scale} to scale the measurements collected from our dimer and record them as $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$. These measurements signals are used to calculate $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$, because we know $\hat{s}$ we can determine which reference orientation is closest, denoted as $\hat{\bf n}_{best}$
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing the Model}
\label{sec:test}

Using our simulation from Section~\ref{sec:brownian} we calculated the
instantaneous signal from our dimer at each time step using MSTM.  We
applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} ---
and the reference orientation with the highest probability is taken as
our estimation of the dimer's instantaneous orientation.  The prior of
the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each
time step by accounting for the physical distance between the previous
estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference
orientation $\hat{\bf n}_\alpha$
\begin{align}
  r_i(t, \hat{\bf n}_\alpha)
  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
  \\
  p_i(\hat{\bf n})
  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
\end{align}
where $\beta$ is a weighting factor that controls the significance of
our priori estimate. The simulation data from Section~\ref{sec:brownian}
was used to check if the estimation was correct by comparing the
estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation
closest to the instantaneous orientation, denoted as
$\hat{\bf n}_{best}$.  To measure the error in the model's estimate
the distribution's of $p(\hat{\bf n}| y(\bar{\bf s}))$ and an
ideal result, $p_ideal$, were compared.  An ideal distribution is one
that is 0 for every $\hat{\bf n}$ apart from the correct result.
\begin{align}
p_{ideal} = 
\begin{cases}
	1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
	0 & \text{anywhere else}
\end{cases}
\end{align}
The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation.  The model's error was
measured by calculating the Kullback-Leibler divergence $K_l$ between
the two distributions
\begin{align}
  K_{l, \#}(p(\hat{\bf n}_\alpha| y_1, y_2, y_3), p_{ideal})
  = p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
\right]
\end{align}
where a more negative value of $K_l$ indicates a poorer estimation of
the dimer's instantaneous orientation.  By summing the divergence of
each measurement across the entire simulation we get an evaluation of
how well the model performed in estimating the dimer's orientation.
To compare between different iterations of our model the we compare
our result of $K_l$ to a worst case scenario and evaluate how much it
improves upon this, denoted as $F(K_l)$:
\begin{align}
K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

A key assumption of Eq.~\eqref{eq:conditional} is that we assume our
detected light signal will inherently have some degree of noise. This
is the central premise that allows us to form
Eq.~\eqref{eq:conditional}, understanding the limits of acceptable
signal error is necessary for implementing this to experimental work.
To this end we ran our model while varying the value of $\epsilon$.
\figurename{ 3} shows the performance of the method with a selection
of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set
to $1$.  The dotted line indicates the minimum radian distance
($0.137$ radians) between two neighbouring reference orientations, if
we are under this line then we know our model is just off from
selecting the correct orientation

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
Section~\ref{sec:Bayes} covered how Bayes theorem can be applied to estimate the orientation of a dimer based on its scattering signal. To evaluate the estimation we can use the simulation in Section~\ref{sec:brownian}; we denote the dimer's \emph{actual} orientation as $\hat{\bf s}$ and use MSTM to calculate its light scattering $I(\hat{\bf s}, \theta)$. At each angle $\theta_k$ we use \eqref{eq:scale} to scale the measurements collected from our dimer and record them as $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$. These measurements signals are used to calculate $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$, because we know $\hat{s}$ we can determine which reference orientation is closest, denoted as $\hat{\bf n}_{best}$.

To measure the error in the model's estimate the distribution from Eq.~\eqref{eq:bayes} and an ideal result, $p_ideal$, were compared.  An ideal result would be one that is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$.
\begin{align}
	p_{ideal} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation.  The model's error was
measured by calculating the Kullback-Leibler divergence $K_l$ between
the two distributions.
\begin{align}
	K_{l, \#}(p_{ideal}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its prediction of the dimer's orientation. The divergence provides us a measure of how confident our measurement is. To better reflect the model's accuracy we implemented a penalty condition that worsens the divergence if our model estimates the wrong reference orientation:
\begin{align}
	p(\hat{\bf n}|y_1,y_2,y_3) = 
	\begin{cases}
		p(\hat{\bf n}_{estimate}|y_1,y_2,y_3) & 
		\text{when $\hat{\bf n}_{estimate}$ = $\hat{\bf n}_{best}$}\\
		p(\hat{\bf n}_{estimate}|y_1,y_2,y_3)-p(\hat{\bf n}_{best}|y_1,y_2,y_3) & \text{when $\hat{\bf n}_{estimate} \ne \hat{\bf n}_{best}$}
	\end{cases}
\end{align}

This allows us to account for situations when our collected signal is equally likely to be associated with 2 reference signals. Section~\ref{sec:Bayes} and Section~\ref{sec:divergence} are all for a single measurement of a light scattering signal from our dimer. To properly model it's motion we need to consider how the orientation changes with time. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing the Model}
\label{sec:test}

Using our simulation from Section~\ref{sec:brownian} we calculated the instantaneous signal from our dimer at each time step using MSTM.  We applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:Bayes} --- and the reference orientation with the highest probability is taken as our estimation of the dimer's instantaneous orientation. To visualise our model's estimation we plotted the distance between our estimation $\hat{\bf n}_{est}$ of the orientation and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ with time. If we assume our prior of the reference orientations $p(\hat{\bf n}_\alpha)$ to be uniform the performance is rather poor, as seen in. 
The prior of the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each time step by accounting for the physical distance between the previous estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$
\begin{align}
  r_i(t, \hat{\bf n}_\alpha)
  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
  \\
  p_i(\hat{\bf n})
  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
\end{align}
where $\beta$ is a weighting factor that controls the significance of our priori estimate. The simulation data from Section~\ref{sec:brownian} was used to check if the estimation was correct by comparing the estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation closest to the instantaneous orientation, denoted as
$\hat{\bf n}_{best}$.  By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation.
To compare between different iterations of our model the we compare our result of $K_l$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
\begin{align}
K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right] \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

A key assumption of Eq.~\eqref{eq:conditional} is that we assume our
detected light signal will inherently have some degree of noise. This
is the central premise that allows us to form
Eq.~\eqref{eq:conditional}, understanding the limits of acceptable
signal error is necessary for implementing this to experimental work.
To this end we ran our model while varying the value of $\epsilon$.
\figurename{ 3} shows the performance of the method with a selection
of angles $15^{\circ}$, $55^{\circ}$, and $90^{\circ}$ and $\beta$ set
to $1$.  The dotted line indicates the minimum radian distance
($0.137$ radians) between two neighbouring reference orientations, if
we are under this line then we know our model is just off from
selecting the correct orientation:
